Visualizations: Metropolitan Areas per Capita, Wave I, II, III (Deaths)

### 2020-11-24 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd
US States per Capita - 3 waves
Counties next

names(msa_region)

region_fips_join <- msa_region %>% 
  full_join(region_fips, by = "agg_index") %>%
  drop_na(fips)

View(msa_covid_only)

View(region_fips_join)
View(covid)

final_daily <- msa_fips_join %>% # or final total 
  bind_rows(region_fips_join) %>% 
  select(index, Date, state, agg_index, fips, csa, county, population, 
         Confirmed, Deaths, new_confirmed, new_deaths, 
         new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
         per_100k_new_confirmed_7, per_100k_new_deaths_7)

names(final_daily)
# write.csv(final_daily, "data/final-50-county-daily.csv")

@@

@@

county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")

@@

# add lat long attributes to county dataset          
county_geo <- merge(county_lines, final_daily, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

map_daily <- county_map # daily covid map

# Wave/Dates Filter  
### For total cases

```{r}
# library(spdplyr) # spatial version of dplyr

# Dates of Waves

# all cases map 
map_total <- county_map %>% 
  filter(Date == max(as.Date(county_map$Date), na.rm = TRUE)) %>% # need to remove NA
  rename(state_fips = STATEFP) %>%
  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)

# wave 2 start date 
map_w2_start <- map_daily %>%
  filter(Date == "2020-06-15") %>% 
  rename(state_fips = STATEFP) %>%
  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)

# wave 2 map 
map_w2 <- map_total %>% 
  st_join(map_w2_start, by = "fips") %>% 
  mutate(Confirmed = Confirmed.x - Confirmed.y) %>% 
  mutate(Deaths = Deaths.x - Deaths.y) %>% 
  rename(start_date = Date.y) %>% 
  rename(end_date = Date.x) %>% 
  rename(fips = fips.x) %>% 
  rename(county = NAME.x) %>% 
  rename(population = population.x) %>% 
  rename(csa = csa.x) %>% 
  rename(agg_index = agg_index.x) %>% 
  rename(state = state.x) %>%
  rename(index = index.x) %>%
  rename(county_fips = COUNTYFP.x) %>%
  rename(state_fips = state_fips.x) %>%
  select(fips, index, agg_index, state, county, csa, population, Confirmed, Deaths, start_date, end_date, state_fips, county_fips, geometry, LSAD.x, ALAND.x, AWATER.x)

@@

```{r}
library(usmap) # to get US Map Data

@@

# county state map 
mapping <- function(map_type, pax) {
  # 
  # fips_code <- fips(state = state_name) 
  # 
  # map <- map_type %>% 
  #   filter(state_fips == fips_code) %>% 
  #   filter(Deaths >= 0)
  #

  map <- map_type

  ggplot(data = map) +
    geom_sf(aes(fill = Deaths, color = Deaths)) + 
    scale_fill_viridis_c(option = "viridis", direction = -1) + 
    labs(title = "something", #paste(state),
         subtitle = "Total Deaths") +
    theme_void() + 
    theme(text = element_text(family = "raleway"))

  # if (pax = FALSE) {
  #   
  #   map <- 
  #   
  # } else if (pax = TRUE) {
  #   
  #   
  #

}

dim(map_w2)

length(map_total$Deaths)
length(map_w2$Deaths)

mapping(map_w2)

>>>>

# problematic with states without fips 
# region_fips_join <- msa_region %>% 
#   full_join(region_fips, by = "agg_index") %>%
#   drop_na(fips)

# final_daily <- msa_fips_join %>% # or final total 
#   bind_rows(region_fips_join) %>% 
#   select(index, Date, state, agg_index, fips, csa, county, population, 
#          Confirmed, Deaths, new_confirmed, new_deaths, 
#          new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
#          per_100k_new_confirmed_7, per_100k_new_deaths_7)
#          
# names(final_daily)
# write.csv(final_daily, "data/final-50-county-daily.csv")

@@

# Getting Wave Count

state_covid_only <- state_covid_only %>% 
  rename("NAME" = "agg_index") %>%
  filter(! NAME %in% c("Alaska", "Hawaii", "Rhode Island", "District of Columbia"))

state_total <- state_covid_only %>% 
  filter(Date == max(as.Date(state_covid_only$Date), na.rm = TRUE)) %>% # need to remove NA
  mutate(Date_Total = Date)

state_w2_start <- state_covid_only %>% 
  filter(Date == max(wave_2)) %>%
  mutate(Date_W2 = Date)

state_w3_start <- state_covid_only %>% 
  filter(Date == max(wave_3)) %>% 
  mutate(Date_W3 = Date)

state_all <- state_total %>% 
  # joining wave 2 and beyond numbers
  left_join(state_w2_start, by = "NAME") %>% 
  rename(Confirmed_total = Confirmed.x) %>%
  rename(Deaths_total = Deaths.x) %>%
  # wave 1 only %>% 
  mutate(Confirmed_W1 = Confirmed.y) %>% 
  mutate(Deaths_W1 = Deaths.y) %>%
  # wave 2 and beyond column 
  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed.y) %>% 
  mutate(Deaths_W2_plus = Deaths_total - Deaths.y) %>% 
  # joining wave 3 numbers 
  left_join(state_w3_start, by = "NAME") %>%
  # only wave 2 
  mutate(Deaths_W2_only = Deaths - Deaths_W2_plus) %>% 
  mutate(Confirmed_W2_only = Confirmed - Confirmed_W2_plus) %>% 
  # wave 3 and beyond
  mutate(Deaths_W3 = Deaths_total - Deaths) %>% 
  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>%
  # tidy up, selecting only relevant variables
  select(index, NAME, population, 
         Date_Total, Confirmed_total, Deaths_total, 
         Confirmed_W1, Deaths_W1,
         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only, 
         Date_W3, Confirmed_W3, Deaths_W3)

View(state_all)
```

@@

county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")
state_lines <- st_read("data/state_shape/cb_2018_us_state_500k.shp")

@@

# add lat long attributes to county dataset          
county_geo <- merge(county_lines, msa_covid_only, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

# merge state 
# add lat long attributes to state dataset

state_geo <- merge(state_lines, state_all, by = "NAME")
state_map <- state_geo %>% st_transform('+proj=longlat +datum=WGS84')

@@

options(scipen = 999)

library(usmap) # to get US Map Data
library(cowplot) # for grid plot

@@

g_total <- ggplot(data = state_map) +
  geom_sf(aes(fill = Deaths_total/population), size = 0.3) +  # color = NA 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
  labs(title = "Total Deaths\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 50))

g_w1 <- ggplot(data = state_map) +
  geom_sf(aes(fill = Deaths_W1/population), size = 0.3) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
  labs(title = "Wave 1 Deaths Only\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 50))

g_w2 <- ggplot(data = state_map) +
  geom_sf(aes(fill = Deaths_W2_only/population), size = 0.3) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
  labs(title = "Wave 2 Deaths Only\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 50))

g_w3 <- ggplot(data = state_map) +
  geom_sf(aes(fill = Deaths_W3/population), size = 0.3) + 
  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
  labs(title = "Wave 3 Deaths Onwards\n") + 
  theme_void() + 
  theme(text = element_text(family = "raleway", size = 50))

p <- plot_grid(g_total, g_w2, g_w3, ncol = 3)

title <- ggdraw() + 
  draw_label("Total Deaths per Capita in U.S. by State per Wave", fontface='bold') + 
  theme(text = element_text(family = "raleway"))

plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))

# ggsave("viz/state-waves.png", width = 8, height = 3, dpi = 300)
ggsave("viz/state-waves-total.png", g_total, width = 8, height = 5, dpi = 300)
ggsave("viz/state-waves-W1.png", g_w1, width = 8, height = 5, dpi = 300)
ggsave("viz/state-waves-W2.png", g_w2, width = 8, height = 5, dpi = 300)
ggsave("viz/state-waves-W3.png", g_w3, width = 8, height = 5, dpi = 300)
